The Chemistry Development Kit (CDK) \cdkversion{}: atom typing, depiction, molecular formulas, and substructure searching

### 2017-04-30 article.tex
Small tweaks

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
The Chemistry Development Kit (CDK) is a 
widely used open source cheminformatics toolkit, providing data
structures to represent chemical concepts along with methods to
manipulate such structures and perform computations on them. The
library implements a wide variety of cheminformatics algorithms
ranging from chemical structure canonicalization to molecular
descriptor calculations and pharmacophore perception. It is used
in drug discovery, metabolomics, and toxicology. Over the
last ten years, the code base has grown significantly resulting in
many complex interdependencies among components and poor
performance of many algorithms. 
\parttitle{Results}
We report improvements to the CDK \cdkversion{} since the v1.2 release series,
specifically addressing the increased functional complexity and poor
performance. We first summarize the addition of new, and
improvements to existing functionality that has led to significant
improvement in performance and support for important chemistry
features including atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
Second, we outline how the CDK has evolved with respect to
quality control and the approaches we have adopted to ensure stability, including
a code review mechanism.
\parttitle{Conclusions}
This paper highlights our continued efforts to provide a community
driven, open source
cheminformatics library, and shows that such collaborative projects can thrive over
extended periods of time, resulting in a high-quality and performant
library. By taking 
advantage of community support and contributions, we show 
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

  Along with the new stereochemistry representation, algorithms were required in several areas. Generally
  a user does not need to invoke these procedures explicitly as they are called as needed within existing
  APIs:

@@

The SMILES parsing has been replaced by code from the external Beam project~\cite{Beam}.
This BSD-licensed SMILES parser is a complete implementation of the SMILES
and OpenSMILES (\url{http://opensmiles.org/}) specifications by one of the
authors (including stereochemistry), and is independent of
the CDK library. The SmilesParser API uses this library underneath, and the
Beam API is hidden by this class. Basic usage is as follows:

@@

The molfile format is still very popular and despite it being a proprietary
format it has become a de facto standard. The format forms the core of the larger
CTfile family which was originally developed by MDL Information Systems~\cite{Dalby92}. The
current format specification is published by BIOVIA and available on 
request~\cite{ctfilespec}.

@@

The CDK library is now mostly written that it no longer depends on a specific
implementation of the \texttt{IChemObjectBuilder}, allowing the user of the
CDK to select a builder suitable to their software. Therefore, if software
depends on event passing, then the \texttt{DefaultChemObjectBuilder} can be
used, in most cases this isn't needed and the \texttt{SilentChemObjectBuilder} 
is preferred resulting in a typical speed up of 10\%-20\%:

@@

The benchmark was run on Java SE 8, Cent OS 7, Intel Core i7-4790 CPU @ 
3.60GHz with 16 GB of RAM. The code to run the benchmark is available in 
the supplementary material allowing numbers to be recorded on the
reader's system.

>>>>

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
The Chemistry Development Kit (CDK) is a 
widely used open source cheminformatics toolkit, providing data
structures to represent chemical concepts along with methods to
manipulate such structures and perform computations on them. The
library implements a wide variety of cheminformatics algorithms
ranging from chemical structure canonicalization to molecular
descriptor calculations and pharmacophore perception. It is used
in drug discovery, metabolomics, and toxicology. Over the
last ten years, the code base has grown significantly, however, resulting in
many complex interdependencies among components and poor
performance of many algorithms. 
\parttitle{Results}
We report improvements to the CDK \cdkversion{} since the v1.2 release series,
specifically addressing the increased functional complexity and poor
performance. We first summarize the addition of new, and
improvements to existing functionality that has led to significant
improvement in performance and support for important chemistry
features including atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
Second, we outline how the CDK has evolved with respect to
quality control and the approaches we have adopted to ensure stability, including
a code review mechanism.
\parttitle{Conclusions}
This paper highlights our continued efforts to provide a community
driven, open source
cheminformatics library, and shows that such collaborative projects can thrive over
extended periods of time, resulting in a high-quality and performant
library. By taking 
advantage of community support and contributions, we show 
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

  Along with the new stereochemistry representation, algorithms were required in several areas. Generally,
  a user does not need to invoke these procedures explicitly as they are called as needed within existing
  APIs:

@@

The SMILES parsing has been replaced by code from the external Beam project~\cite{Beam,Weininger1988}.
This BSD-licensed SMILES parser is a complete implementation of the SMILES
and OpenSMILES (\url{http://opensmiles.org/}) specifications by one of the
authors (including stereochemistry), and is independent of
the CDK library. The SmilesParser API uses this library underneath, and the
Beam API is hidden by this class. Basic usage is as follows:

@@

The molfile format is still very popular and despite it being a proprietary
format, it has become a de facto standard. The format forms the core of the larger
CTfile family which was originally developed by MDL Information Systems~\cite{Dalby92}. The
current format specification is published by BIOVIA and available on 
request~\cite{ctfilespec}.

@@

The CDK library is now mostly written that it no longer depends on a specific
implementation of the \texttt{IChemObjectBuilder}, allowing the user of the
CDK to select a builder suitable to their software. Therefore, if software
depends on event passing, then the \texttt{DefaultChemObjectBuilder} can be
used, in most cases this isn't needed and the \texttt{SilentChemObjectBuilder} 
is preferred resulting in a typical speed up of 10\% to 20\%:

@@

The benchmark was run on Java SE 8, CentOS 7, Intel Core i7-4790 CPU @ 
3.60GHz with 16 GB of RAM. The code to run the benchmark is available in 
the supplementary material allowing numbers to be recorded on the
reader's system.
